---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.retina = 3
)
```

# ggsegFreeSurfer

<!-- badges: start -->
[![R-CMD-check](https://github.com/ggsegverse/ggsegFreeSurfer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ggsegverse/ggsegFreeSurfer/actions/workflows/R-CMD-check.yaml)
[![r-universe](https://ggseg.r-universe.dev/badges/ggsegFreeSurfer)](https://ggseg.r-universe.dev/ggsegFreeSurfer)
<!-- badges: end -->

This package provides FreeSurfer cortical and subcortical brain atlases
formatted for use with ggseg.

## Installation

We recommend installing the ggseg-atlases through the ggseg
[r-universe](https://ggseg.r-universe.dev/ui#builds):

```{r}
#| label: install-r-universe
#| eval: false
options(repos = c(
  ggseg = "https://ggseg.r-universe.dev",
  CRAN = "https://cloud.r-project.org"
))

install.packages("ggsegFreeSurfer")
```

You can install this package from [GitHub](https://github.com/) with:

```r
# install.packages("pak")
pak::pak("ggsegverse/ggsegFreeSurfer")
```

# Cortical atlases

## Desikan-Killiany-Tourville (DKT) atlas

```{r}
#| label: dkt
library(ggseg)
library(ggsegFreeSurfer)

plot(dkt())
```

## Destrieux atlas

```{r}
#| label: destrieux
#| fig-height: 7
plot(destrieux())
```

# Subcortical atlases

The subcortical atlases are embedded in the full-brain `aseg`, so each slice
view shows the structure of interest within its anatomical context (cerebral
cortex and white matter rendered grey).

## Anterior/posterior hippocampus atlas

```{r}
#| label: hcpa
#| fig-height: 4
plot(hcpa())
```

## Thalamic nuclei atlas

```{r}
#| label: thalamus
#| fig-height: 5
plot(thalamus())
```

## Hippocampal subfields & amygdala nuclei atlas

```{r}
#| label: hippoamyg
#| fig-height: 5
plot(hippoamyg())
```

## Brainstem substructures atlas

```{r}
#| label: brainstem
#| fig-height: 4
plot(brainstem())
```

## Hypothalamic subunits atlas

```{r}
#| label: hypothalamus
#| fig-height: 4
plot(hypothalamus())
```

## Data source

Desikan RS, Segonne F, Fischl B, Quinn BT, Dickerson BC, Blacker D, ... &
Killiany RJ (2006). An automated labeling system for subdividing the human
cerebral cortex on MRI scans into gyral based regions of interest. *NeuroImage*,
31(3), 968-980.

Destrieux C, Fischl B, Dale A, & Halgren E (2010). Automatic parcellation of
human cortical gyri and sulci using standard anatomical nomenclature.
*NeuroImage*, 53(1), 1-15.

Iglesias JE, Augustinack JC, Nguyen K, Player CM, Player A, Wright M, ... &
Van Leemput K (2015). A computational atlas of the hippocampal formation using
ex vivo, ultra-high resolution MRI: Application to adaptive segmentation of in
vivo MRI. *NeuroImage*, 115, 117-137.

Iglesias JE, Van Leemput K, Bhatt P, Casillas C, Dutt S, Schuff N, ... &
Fischl B (2015). Bayesian segmentation of brainstem structures in MRI.
*NeuroImage*, 113, 184-195.

Saygin ZM, Kliemann D, Iglesias JE, van der Kouwe AJW, Boyd E, Reuter M, ... &
Fischl B (2017). High-resolution magnetic resonance imaging reveals nuclei of
the human amygdala: manual segmentation to automatic atlas. *NeuroImage*, 155,
370-382.

Iglesias JE, Insausti R, Lerma-Usabiaga G, Bocchetta M, Van Leemput K, Greve DN,
... & Paz-Alonso PM (2018). A probabilistic atlas of the human thalamic nuclei
combining ex vivo MRI and histology. *NeuroImage*, 183, 314-326.

Billot B, Bocchetta M, Todd E, Dalca AV, Rohrer JD, & Iglesias JE (2020).
Automated segmentation of the hypothalamus and associated subunits in brain MRI.
*NeuroImage*, 223, 117287.
