---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.retina = 3
)
```

# ggsegFreeSurfer

<!-- badges: start -->
[![R-CMD-check](https://github.com/ggsegverse/ggsegFreeSurfer/workflows/R-CMD-check/badge.svg)](https://github.com/ggsegverse/ggsegFreeSurfer/actions)
<!-- badges: end -->

This package provides FreeSurfer cortical and subcortical brain atlases
formatted for use with
[ggseg](https://ggseg.github.io/ggseg/) and
[ggseg3d](https://ggseg.github.io/ggseg3d/).

| Atlas | Function | Type | Regions/hemi |
|-------|----------|------|--------------|
| Desikan-Killiany-Tourville | `dkt()` | Cortical | 32 |
| Destrieux | `destrieux()` | Cortical | 75 |
| HCP Subcortical | `hcpa()` | Subcortical | -- |

To learn how to use these atlases, please look at the documentation for
[ggseg](https://ggseg.github.io/ggseg/) and
[ggseg3d](https://ggseg.github.io/ggseg3d/).

## Installation

You can install ggsegFreeSurfer from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("ggsegverse/ggsegFreeSurfer")
```

## Atlases

```{r}
#| label: dkt
#| fig-height: 4
library(ggseg)
library(ggsegFreeSurfer)
library(ggplot2)

ggplot() +
  geom_brain(
    atlas = dkt(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Desikan-Killiany-Tourville") +
  theme_void()
```

```{r}
#| label: destrieux
#| fig-height: 4
ggplot() +
  geom_brain(
    atlas = destrieux(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Destrieux (aparc.a2009s)") +
  theme_void()
```

```{r}
#| label: hcpa
library(ggseg3d)

ggseg3d(atlas = hcpa())
```
