---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.retina = 3
)
```

# ggsegFreeSurfer

<!-- badges: start -->
[![R-CMD-check](https://github.com/ggsegverse/ggsegFreeSurfer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ggsegverse/ggsegFreeSurfer/actions/workflows/R-CMD-check.yaml)
[![r-universe](https://ggseg.r-universe.dev/badges/ggsegFreeSurfer)](https://ggseg.r-universe.dev/ggsegFreeSurfer)
<!-- badges: end -->

This package provides FreeSurfer cortical and subcortical brain atlases
formatted for use with ggseg.

## Installation

We recommend installing the ggseg-atlases through the ggseg
[r-universe](https://ggseg.r-universe.dev/ui#builds):

```{r}
#| label: install-r-universe
#| eval: false
options(repos = c(
  ggseg = "https://ggseg.r-universe.dev",
  CRAN = "https://cloud.r-project.org"
))

install.packages("ggsegFreeSurfer")
```

You can install this package from [GitHub](https://github.com/) with:

```r
# install.packages("pak")
pak::pak("ggsegverse/ggsegFreeSurfer")
```

## Desikan-Killiany-Tourville (DKT) atlas

```{r}
#| label: dkt
library(ggseg)
library(ggsegFreeSurfer)
library(ggplot2)

ggplot() +
  geom_brain(
    atlas = dkt(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  scale_fill_manual(values = dkt()$palette, na.value = "grey") +
  theme_void()
```

## Destrieux atlas

```{r}
#| label: destrieux
#| fig-height: 7
ggplot() +
  geom_brain(
    atlas = destrieux(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  scale_fill_manual(values = destrieux()$palette, na.value = "grey") +
  theme_void()
```

## Data source

Desikan RS, Segonne F, Fischl B, Quinn BT, Dickerson BC, Blacker D, ... &
Killiany RJ (2006). An automated labeling system for subdividing the human
cerebral cortex on MRI scans into gyral based regions of interest. *NeuroImage*,
31(3), 968-980.

Destrieux C, Fischl B, Dale A, & Halgren E (2010). Automatic parcellation of
human cortical gyri and sulci using standard anatomical nomenclature.
*NeuroImage*, 53(1), 1-15.
